### 2016-12-07 Projects/DelCo_2012/Report/DelCo_2012.tex
Minor edits to CAF report

Twenty-one of the spray density tests were conducted in the concrete burn building that was later used for the gas cooling tests, and twelve of the tests were conducted in the experimental structures that were later used for the fire suppression tests. The test variables were the nozzle height, nozzle location within the test structure, and the nozzle pattern (Figure~\ref{fig:Spray_Density_Nozzle_Patterns}). Descriptions of these facilities and their floor plan schematics are included in Section~\ref{sec:Experimental Structures}. A detailed discussion of the spray density tests and results is included in Section~\ref{sec:Spray_Density}.

@@

The structures were instrumented for gas temperature, gas velocity, and heat flux measurements. Gas temperatures in the burn rooms were measured with bare-bead, Chromel-Alumel (type K) thermocouples. Additional single, sheathed thermocouples were installed in conjunction with the bi-directional probes for gas velocity measurements. The single thermocouples were bare-bead, Chromel-Alumel (type K) thermocouples with a 1.0~mm (0.04~in) nominal diameter. The thermocouple was protected with an 3.2~mm (0.13~in) diameter inconel sheath. The sheathing protects the thermocouple from water but slows the response time because of the additional thermal mass. Schmidt-Boelter gauges were used to measure both total heat flux and radiant heat flux (radiometer). A radiometer is a total heat flux gauge with a zirconium plate to prevent contributions from convective heat transfer. A legend, which clarifies the instrumentation schematics discussed in the following sections, is included in Figure~\ref{fig:Instrumentation_Legend}.

@@

The gas cooling tests included four bare-bead thermocouple arrays, a bi-directional probe with a thermocouple array, and a total heat flux gauge/radiometer pair. Figure~\ref{fig:Gas_Cooling_Instrumentation_Dimensions} provides the positions within the burn room and gas cooling room where the sensors were located (see Figure~\ref{fig:Instrumentation_Legend} for reference on the symbols). Each of the bare-bead thermocouple arrays had 11 thermocouples. The top thermocouple was placed at the ceiling and each subsequent thermocouple was spaced 0.3~m (1~ft) apart with the bottom thermocouple being 3.05~m (10~ft) below the ceiling. There were six velocity probes and solid thermocouples centered at the external doorway to the gas cooling room. The top probe was 0.15~m (0.5~ft) below the door soffit, the second probe was 0.3~m (1~ft) below the soffit, and the remaining four were spaced 0.3~m (1~ft) apart with the bottom probe being 1.52~m (5~ft) below the soffit. The total heat flux gauge/radiometer set was positioned 0.15~m (0.5~ft) off the ground and aimed to view the ceiling.

@@

The fire suppression testing in the single-story structure included two bare-bead thermocouple arrays, three bi-directional probe plus solid thermocouple arrays, two total heat flux gauge pairs, and two total heat flux gauge/radiometer pairs. Figure~\ref{fig:Fire_Suppression_Instrumentation_Dimensions} provides the positions within the burn room, hallway, and entrance foyer where the sensors were located (see Figure~\ref{fig:Instrumentation_Legend} for reference on the symbols). Each bare-bead thermocouple array had eight thermocouples. The top sensor was 0.03~m (1~in) below the ceiling and the remaining thermocouples were spaced 0.3~m (1~ft) apart with the bottom thermocouple being 2.13~m (7~ft) below the ceiling. The three bi-directional probe arrays had unique sensor locations. For the window array, there were four sensor pairs located 0.28~m (0.95~ft), 0.58~m (1.90~ft), 0.84~m (2.85~ft), and 1.16~m (3.8~ft) below the soffit. For the hallway array, the seven sensor pairs started 0.3~m (1~ft) below the ceiling and were spaced every 0.3~m (1~ft) ending 2.13~m (7~ft) below the ceiling. The doorway array also had seven sensor pairs, but the first pair was located 0.15~m (0.5~ft) below the soffit. The remaining sensors were spaced every 0.3~m (1~ft) with the bottom pair 1.83~m (6~ft) below the doorway soffit. The total heat flux gauge/radiometer pairs were set to be 0.15~m (0.5~ft) off the ground and aimed to view the ceiling. The pairs of total heat flux gauges were set 1~m (3~ft) off the ground with one sensor ``looking'' vertical at the ceiling and the other horizontal.

@@

Fire suppression testing in the two-story structure included similar types of instrumentation but with more sensors. Three bare-bead thermocouple arrays and two bi-directional probe and thermocouple arrays were used in the first floor. Figure~\ref{fig:fire_supp_first_2story} shows the positions of the instrumentation (see Figure~\ref{fig:Instrumentation_Legend} for reference on the symbols). Each of the bare-bead arrays had eight thermocouples with the same spacing as the single-story tests. The bi-directional probe and solid thermocouple arrays also had eight sensors with the first probe 0.08~m (0.25~ft) below the soffit and the remaining probes at 0.34~m (1.1~ft), 0.61~m (2~ft), 0.88~m (2.9~ft), 1.15~m (3.7~ft), 1.42~m (4.7~ft), 1.68~m (5.5~ft), and 1.95~m (6.4~ft) below the soffit.

@@

The second story featured six bare-bead arrays and three bi-directional probe plus solid thermocouple arrays. The second floor also had three heat flux sensor pairs (one facing horizontal and one facing vertical) that were 1~m (3~ft) off the ground. Figure~\ref{fig:fire_supp_second_2story} shows the positions of the instrumentation (see Figure~\ref{fig:Instrumentation_Legend} for reference on the symbols). The bare-bead thermocouple arrays featured eight sensors with the same spacing as the basement level. The bi-directional probe arrays at the south wall door and at the top of the stairwell were comprised of eight sensor pairs with the same spacing as the ground level. The third bi-directional probe array (3~m (9~ft) off the north wall and 0.6~m (2~ft) off the east wall) had eight sensor pairs with the same spacing at the bare-bead thermocouple arrays.

@@

There are different components of uncertainty in the length, mass, temperature, heat flux, gas concentration, differential pressure, gas velocity, and heat release rate reported here. Uncertainties are grouped into two categories according to the method used to estimate them. Type A uncertainties are those which are evaluated by statistical methods, and Type B are those which are evaluated by other means~\cite{Taylor&Kuyatt:1994}. Type B analysis of systematic uncertainties involves estimating the upper (+~a) and lower (-~a) limits for the quantity in question such that the probability that the value would be in the interval ($\pm$~a) is essentially 100~\%. After estimating uncertainties by either Type A or B analysis, the uncertainties are combined in quadrature to yield the combined standard uncertainty. Then, the combined standard uncertainty is multiplied by a coverage factor of two, which results in the expanded uncertainty with a 95~\% confidence interval (2$\sigma$).  For some of these components, such as the zero and calibration elements, uncertainties are derived from referenced instrument specifications. For other components, referenced research results and past experience with the instruments provided input in the uncertainty determination.

@@

The gas measurement instruments and sampling system used in this series of experiments have demonstrated an expanded (k~=~2) relative uncertainty of $\pm$~1~\% when compared with span gas volume fractions~\cite{Bundy:2007}. Given the non-uniformities and movement of the fire gas environment and the limited set of sampling points in these experiments, an estimated uncertainty of $\pm$~12~\% was applied to the results~\cite{Lock:1}.

@@

Bi-directional probes and single thermocouples were used to measure gas velocity. The bi-directional probes used similar pressure transducers as those used for the differential pressure measurements discussed above. Bare-bead Type K thermocouples are co-located with each probe. A gas velocity measurement study examining the doorway flow of pre-flashover compartment fires yielded expanded uncertainty measurements ranging from $\pm$~0.14 to $\pm$~0.22 for bi-directional probes of similar design~\cite{Bryant:FSJ2009}. The total expanded uncertainty for gas velocity in these experiments is estimated to be $\pm$~18~\%.

Water flow rate was measured with a pressure and flow meter combination. The meter consists of a section of 63.5~mm (2.5~in) cast aluminum pipe with a 0--4.1~MPa (0--600~psi) pressure transducer and a paddlewheel type flow sensor with a range of 0 to 4800~l/min (1250~gpm). The pressure transducer and paddlewheel both connect to the battery operated control box where the pressure transducer voltage is converted to a pressure and the paddlewheel pulse count is converted to a volumetric flow rate.  The manufacturer reports a $\pm$~5~\% calibration expanded uncertainty for the flow sensor and $\pm$~3~\% calibration expanded uncertainty for the pressure sensor~\cite{Akron:2009}. The pressure transducer was calibrated with a known analog pressure gauge. The flow meter was calibrated by capturing water over time and measuring that mass of water to determine the flow rate. The total expanded uncertainty is estimated at $\pm$~10~\%.

@@

Wood pallets were used as the fuel source for the gas cooling experiments. The pallets were approximately 1.2~m (4.0~ft) by 1.0~m (3.3~ft) by 0.13~m (0.42~ft) thick and ranged in mass from 13.6~kg (29.9~lb) to 26.4~kg (58.1~lb) with an average of 18.4~kg (40.5~lb). The initial fuel load consisted of 10 pallets, arranged in two stacks of five, as shown in Figure~\ref{fig:Burn_Building_Fuel_Load}. Approximately one half of a bale of excelsior, 13.0~kg (28.6~lb), was mixed with the pallets to aid with ignition. Figure~\ref{fig:Burn_Building_Fuel_Load} also provides dimensioned locations of the two stacks of pallets within the burn room.

@@

Note that as the pallets burned away and the hot gas layer temperatures decreased, the steel shutters on the window to the burn room were opened and additional pallets were added. Pallets were added to the existing fuel locations until the flames from the piles reached the ceiling of the burn room again. Then, the shutters on the window were closed, and the temperature of the hot gas layer in the adjacent room was monitored in preparation for starting another gas cooling experiment.  The number of pallets added each time varied.

@@

The fuel load for test numbers 1 through 8 consisted of two sleeper sofas, two cushioned chairs, carpet, padding, and wood paneling along the southeast corner walls. An image of the fuel arrangement is shown in Figure~\ref{fig:Furniture_Fuel_Load}. The sofas and chairs were set up to mimic a typical seating area with one couch against the south wall and one couch against the east wall, each positioned 0.9~m (36~in) away from the southeast corner. A chair was placed adjacent to each couch. The specific locations can be found in the dimensioned schematic, Figure~\ref{fig:Furniture_Fuel_Load_Dimensions}. In total, this fuel package contains approximately 336~kg (739~lb) of fuel consisting of wood, flexible polyurethane (PU), polyester (PE),and polypropylene (PP). Table~\ref{tab:Fire_Suppression_Fuel_Masses} provides a breakdown on the fuel composition for these experiments.

@@

The fuel load for test numbers 9 through 12 consisted of one bale of hay, 12 pallets, 11.5 sheets of plywood on the walls, and the wood ``floor assembly''. Two stacks of pallets with hay were used as the first items ignited in each experiment. Each stack was composed of six pallets with a half bale of hay layered between the six pallets. Figure~\ref{fig:Wood_Fuel_Load} shows the pallets prior to ignition. The pallets and hay were weighed before each experiment, and the average mass of the two stacks of pallets and the bale of hay over the four experiments was 232.4~kg (511.3~lb).

@@

The pallet stacks were located 0.3~m (12~in) away from the east and south walls of the burn room (see Figure~\ref{fig:Wood_Fuel_Load_Dimensions}). The pallets were 1.2~m (48~in) long, 1.0~m (40~in) wide, and 123~mm (4.8~in) high. The stacks were spaced 150~mm (6~in) apart. One layer of 13~mm (0.5~in) thick gypsum board panels were laid on the concrete floor under the wood pallets to form a protective layer to minimize thermal damage to the concrete floor. The east and south walls of the burn room were covered with 15~mm (0.59~in) thick sheets of plywood.  The wood ``flooring assembly'', composed of 12 TJIs and nine sheets of OSB, served as the ceiling of the burn room. The open doorway to the burn room on the south side was covered with a 2.4~m (8~ft) tall by 1.2~m (4~ft) wide piece of medium density fiberboard paneling that was 5~mm (0.19~in) thick.

@@

In June of 2015, the larger compartment experiments, experiments 13 through 20, were conducted in the two-story structure.  Two different fuel packages located on the ground level were utilized throughout the experiments: experiments 13 and 14 used a fuel package composed primarily of furniture, and experiments 15 through 20 used the same fuel package with an addition of two stacks of wood pallets.

@@

The fuel load for experiments 13 and 14 consisted of three sofas with an average mass of 48.7~kg; 51.4~m$^2$ (553~ft$^2$) of carpet and padding; and approximately 38 sheets of OSB paneling with dimensions 16~mm (0.63~in) x 1.2~m (4~ft) x 2.4~m (8~ft) along the east, west, and south walls and ceiling. The three couches were aligned along the west wall and spaced 1.2~m (4~ft) apart. The specific locations can be found in the dimensioned schematic Figure~\ref{fig:furniture_2story}. In total, this fuel package was composed of wood, flexible polyurethane (PU), polyester (PE), and polypropylene (PP).

@@

For experiments 15 through 20, two stacks of wood pallets were added to the fuel load consisting of three sofas; OSB lining on the ceiling and walls; and the polypropylene carpeting over polyurethane foam padding. Each stack had 10 pallets and an approximate mass of 175~kg (386~lb). Hay from bales with an average a mass of 12~kg (26~lb) was added for ignition purposes. The specific locations of the fuel load's contents can be found in the dimensioned schematic Figure~\ref{fig:pallet_furniture_2story}.

@@

Spray density experiments were conducted by placing 0.23~m$^2$ (2.5~ft$^2$) interlocking water collection pans across the floor surface area of the burn rooms within each structure (see Figure~\ref{fig:water_buckets}). The depth of the pans was 0.3 m (1 ft).  Water or CAF was flowed for a specified duration or until it appeared that a pan was filled.  Upon completion of the test, the bins were removed from the structure. Then, each bin was weighed to quantify the impact of nozzle location and spray pattern. Thirty-three spray density tests were conducted: 21 in the burn building used for the gas cooling tests and 12 in the single-story structure used for the fire suppression experiments. The interior of the fire room in the single-story structure was sheathed in plastic for the spray density tests to protect the gypsum board from being soaked with water prior to the fire suppression experiments. The flow rate for each test was 120~gpm for water and 120~gpm plus 60~cfm air for CAF. Tables~\ref{tab:spray_density_tests} and \ref{tab:spray_density_tests2} provide an overview of each experiment. In the table, ``GC'' stands for the gas cooling experiments, and ``SE'' stands for the suppression experiments. For all experiments, 30~m (100~ft) of 4.45~cm (1.75~in) diameter hoseline connected to a ``blitz-fire'' monitor with a Metro 1 nozzle was used.

@@

For the spray density tests listed in Table~\ref{tab:spray_density_tests}, the ``mid'' position had the centerline of the stream aimed at 2.44~m (8~ft) south of the north wall and 2.13~m (7~ft) west of the east wall in the concrete burn building.  The ``back'' position had the centerline of the stream aimed at 3.36~m (11~ft) south of the north wall and 2.13~m (7~ft) west of the east wall in the concrete burn building. Figure~\ref{fig:stream_position} shows the streams from the monitor nozzle for the middle and back positions. From the images, the qualitative difference of the nozzle positions can be observed.  When the stream is in the middle position, after the stream hits the ceiling, the impact region is relatively narrow with the majority of the water hitting the walls of the corner opposite the nozzle position.  When the nozzle is in the back position, the stream hits the ceiling at a much steeper angle and the droplets from the broken stream have a broader distribution throughout the room.

@@

\begin{table}[!ht]
\footnotesize
\centering
\captionof{table}{Spray Density Test Matrix for Gas Cooling Experiments}\label{tab:spray_density_tests}
\begin{tabular}{lllll}
\toprule[1.5pt]
Test \#    &  Duration (s)  & Agent  &  Pattern            	& Nozzle Position  \\
\midrule
GC1        &  15            & Water  &  Straight Stream    	& Back             \\
GC2        &  23            & Water  &  Straight Stream    	& Back             \\
GC3        &  15            & Water  &  Straight Stream    	& Mid              \\
GC4        &  15            & Water  &  Straight Stream    	& Mid              \\
GC5        &  20            & Water  &  Straight Stream 	& Mid              \\
GC6        &  21            & Water  &  Narrow Fog   		& Mid              \\
GC7        &  20            & Water  &  Narrow Fog   		& Mid              \\
GC8        &  22            & Water  &  Narrow Fog   		& Back             \\
GC9        &  24            & Water  &  Narrow Fog   		& Back             \\
GC10       &  20            & Water  &  Solid Stream       	& Back             \\
GC11       &  22            & Water  &  Solid Stream       	& Back             \\
GC12       &  15            & CAF    &  Solid Stream       	& Back             \\
GC13       &  10            & CAF    &  Solid Stream       	& Mid              \\
GC14       &  15            & CAF    &  Solid Stream       	& Back             \\
GC15       &  20            & CAF    &  Solid Stream       	& Mid              \\
GC16       &  15            & CAF    &  Narrow Fog   		& Mid              \\
GC17       &  20            & CAF    &  Straight Stream    	& Mid             \\
GC18       &  10            & CAF    &  Straight Stream    	& Mid             \\
GC19       &  15            & CAF    &  Narrow Fog   		& Back             \\
GC20       &  15            & CAF    &  Narrow Fog   		& Back             \\
GC21       &  15            & CAF    &  Straight Stream    	& Back             \\
\bottomrule[1.25pt]
\end{tabular}\par
\end{table}
For the fire suppression spray density experiments, Table~\ref{tab:spray_density_tests2}, the hallway position had the center of the stream aimed at the ceiling of the burn room, 2.14~m (7~ft) north of the south wall and 1.53~m (5~ft) east of the west wall. The fire room position had the center of the stream aimed at the ceiling of the burn room, 2.14~m (7~ft) north of the south wall and 3.05~m (10~ft) west of the east wall.

@@

To further analyze the data, the temperature differences can be separated by each of the three thermocouple arrays in the gas cooling room. Figure~\ref{fig:Gas_Cooling_Instrumentation_Dimensions} shows the placement of three thermocouple arrays (TC A1, TC A2, and TC A3). Ideally, the thermocouple is providing an estimate of the temperature of the gas that surrounded it. However, the thermocouple is actually measuring its own temperature. So, if a droplet of water or some CAF clings to a thermocouple, the temperature provided by that thermocouple will be different than the surrounding gas temperature.

>>>>

Twenty-one of the spray density tests were conducted in the concrete burn building that was later used for the gas cooling tests, and 12 of the tests were conducted in the experimental structures that were later used for the fire suppression tests. The test variables were the nozzle height, nozzle location within the test structure, and the nozzle pattern (Figure~\ref{fig:Spray_Density_Nozzle_Patterns}). Descriptions of these facilities and their floor plan schematics are included in Section~\ref{sec:Experimental Structures}. A detailed discussion of the spray density tests and results is included in Section~\ref{sec:Spray_Density}.

@@

The structures were instrumented for gas temperature, gas velocity, and heat flux measurements. Gas temperatures in the burn rooms were measured with bare-bead, Chromel-Alumel (type K) thermocouples. Additional single, sheathed thermocouples were installed in conjunction with the bi-directional probes for gas velocity measurements. The single thermocouples were bare-bead, Chromel-Alumel (type K) thermocouples with a 1.0~mm (0.04~in) nominal diameter. The thermocouple was protected with an 3.2~mm (0.13~in) diameter inconel sheath. The sheathing protects the thermocouple from water but slows the response time because of the additional thermal mass. Schmidt-Boelter gauges were used to measure both total heat flux and radiant heat flux (radiometer). A radiometer is a total heat flux gauge with a zirconium plate to prevent contributions from convective heat transfer. A legend to clarify the instrumentation schematics presented in the following sections is included in Figure~\ref{fig:Instrumentation_Legend}.

@@

The gas cooling tests included four bare-bead thermocouple arrays, a bi-directional probe with a thermocouple array, and a total heat flux gauge/radiometer pair. Figure~\ref{fig:Gas_Cooling_Instrumentation_Dimensions} provides the positions within the burn room and gas cooling room where the sensors were located (see Figure~\ref{fig:Instrumentation_Legend} for reference on the symbols). Each of the bare-bead thermocouple arrays had 11 thermocouples. The top thermocouple was placed at the ceiling and each subsequent thermocouple was spaced 0.3~m (1.0~ft) apart with the bottom thermocouple being 3.1~m (10.0~ft) below the ceiling. There were six velocity probes and solid thermocouples centered at the external doorway to the gas cooling room. The top probe was 0.2~m (0.5~ft) below the door soffit, the second probe was 0.3~m (1.0~ft) below the soffit, and the remaining four were spaced 0.3~m (1.0~ft) apart with the bottom probe being 1.5~m (5.0~ft) below the soffit. The total heat flux gauge/radiometer set was positioned 0.2~m (0.5~ft) off the ground and aimed to view the ceiling.

@@

The fire suppression testing in the single-story structure included two bare-bead thermocouple arrays, three bi-directional probe plus solid thermocouple arrays, two total heat flux gauge pairs, and two total heat flux gauge/radiometer pairs. Figure~\ref{fig:Fire_Suppression_Instrumentation_Dimensions} provides the positions within the burn room, hallway, and entrance foyer where the sensors were located (see Figure~\ref{fig:Instrumentation_Legend} for reference on the symbols). Each bare-bead thermocouple array had eight thermocouples. The top sensor was 0.03~m (1.0~in) below the ceiling and the remaining thermocouples were spaced 0.3~m (1.0~ft) apart with the bottom thermocouple being 2.1~m (7.0~ft) below the ceiling. The three bi-directional probe arrays had unique sensor locations. For the window array, there were four sensor pairs located 0.3~m (1.0~ft), 0.6~m (1.9~ft), 0.9~m (2.9~ft), and 1.2~m (3.8~ft) below the soffit. For the hallway array, the seven sensor pairs started 0.3~m (1.0~ft) below the ceiling and were spaced every 0.3~m (1.0~ft) ending 2.1~m (7.0~ft) below the ceiling. The doorway array also had seven sensor pairs, but the first pair was located 0.2~m (0.5~ft) below the soffit. The remaining sensors were spaced every 0.3~m (1.0~ft) with the bottom pair 1.8~m (6.0~ft) below the doorway soffit. The total heat flux gauge/radiometer pairs were set to be 0.2~m (0.5~ft) off the ground and aimed to view the ceiling. The pairs of total heat flux gauges were set 1.0~m (3.3~ft) off the ground with one sensor ``looking'' vertical at the ceiling and the other horizontal.

@@

Fire suppression testing in the two-story structure included similar types of instrumentation but with more sensors. Three bare-bead thermocouple arrays and two bi-directional probe and thermocouple arrays were used in the first floor. Figure~\ref{fig:fire_supp_first_2story} shows the positions of the instrumentation (see Figure~\ref{fig:Instrumentation_Legend} for reference on the symbols). Each of the bare-bead arrays had eight thermocouples with the same spacing as the single-story tests. The bi-directional probe and solid thermocouple arrays also had eight sensors with the first pair 0.1~m (0.3~ft) below the soffit and the remaining pairs at 0.3~m (1.1~ft), 0.6~m (2.0~ft), 0.9~m (2.9~ft), 1.2~m (3.8~ft), 1.4~m (4.7~ft), 1.7~m (5.5~ft), and 2.0~m (6.4~ft) below the soffit.

@@

The second story featured six bare-bead thermocouple arrays and three bi-directional probe plus solid thermocouple arrays. The second floor also had three heat flux sensor pairs (one facing horizontal and one facing vertical) that were 1.0~m (3.3~ft) off the ground. Figure~\ref{fig:fire_supp_second_2story} shows the positions of the instrumentation (see Figure~\ref{fig:Instrumentation_Legend} for reference on the symbols). The bare-bead thermocouple arrays featured eight sensors with the same spacing as the ground level. The bi-directional probe arrays at the south exterior door and at the top of the stairwell were comprised of eight sensor pairs with the same spacing as the ground level. The third bi-directional probe array located 3.0~m (9.8~ft) off the north wall and 0.6~m (2.0~ft) off the east wall had eight sensor pairs with the same spacing as the bare-bead thermocouple arrays.

@@

There are different components of uncertainty in the length, mass, temperature, heat flux, gas concentration, differential pressure, gas velocity, and heat release rate reported here. Uncertainties are grouped into two categories according to the method used to estimate them. Type A uncertainties are those evaluated by statistical methods, and Type B are those evaluated by other means~\cite{Taylor&Kuyatt:1994}. Type B analysis of systematic uncertainties involves estimating the upper (+~a) and lower (-~a) limits for the quantity in question such that the probability that the value would be in the interval ($\pm$~a) is essentially 100~\%. After estimating uncertainties by either Type A or B analysis, the uncertainties are combined in quadrature to yield the combined standard uncertainty. Then, the combined standard uncertainty is multiplied by a coverage factor of two, which results in an expanded uncertainty with a 95~\% confidence interval (2$\sigma$). For some of these components, such as the zero and calibration elements, uncertainties are derived from referenced instrument specifications. For other components, referenced research results and past experience with the instruments provided input in the uncertainty determination.

@@

The gas measurement instruments and sampling system used in this series of experiments have demonstrated an expanded (k~=~2) relative uncertainty of $\pm$~1~\% when compared with span gas volume fractions~\cite{Bundy:2007}. Given the non-uniformities and movement of the fire gas environment and the limited set of sampling points in these experiments, an estimated uncertainty of $\pm$~12~\% is applied to the results~\cite{Lock:1}.

@@

Bi-directional probes and single thermocouples were used to measure gas velocity. The bi-directional probes used similar pressure transducers as those used for the differential pressure measurements discussed above. Bare-bead Type K thermocouples were co-located with each probe. A gas velocity measurement study examining the doorway flow of pre-flashover compartment fires yielded expanded uncertainty measurements ranging from $\pm$~0.14 to $\pm$~0.22 for bi-directional probes of similar design~\cite{Bryant:FSJ2009}. The total expanded uncertainty for gas velocity in these experiments is estimated to be $\pm$~18~\%.

Water flow rate was measured with a pressure and flow meter combination. The meter consisted of a section of 64~mm (2.5~in) cast aluminum pipe with a 0--4.1~MPa (0--600~psi) pressure transducer and a paddlewheel type flow sensor with a range of 0 to 4730~l/min (1250~gpm). The pressure transducer and paddlewheel both connected to the battery operated control box where the pressure transducer voltage was converted to a pressure and the paddlewheel pulse count was converted to a volumetric flow rate.  The manufacturer reports a $\pm$~5~\% calibration expanded uncertainty for the flow sensor and $\pm$~3~\% calibration expanded uncertainty for the pressure sensor~\cite{Akron:2009}. The pressure transducer was calibrated with a known analog pressure gauge. The flow meter was calibrated by capturing water over time and measuring that mass of water to determine the flow rate. The total expanded uncertainty is estimated at $\pm$~10~\%.

@@

Wood pallets were used as the fuel source for the gas cooling experiments. The pallets were approximately 1.2~m (3.9~ft) by 1.0~m (3.3~ft) by 123~mm (4.8~in) thick and ranged in mass from 13.6~kg (29.9~lb) to 26.4~kg (58.1~lb) with an average of 18.4~kg (40.5~lb). The initial fuel load consisted of 10 pallets, arranged in two stacks of five, at the locations shown in Figure~\ref{fig:Burn_Building_Fuel_Load}. Approximately one half of a bale of excelsior, 13.0~kg (28.6~lb), was mixed with the pallets to aid with ignition.

@@

Note that as the pallets burned away and the hot gas layer temperatures decreased, the steel shutters on the window to the burn room were opened and additional pallets were added. Pallets were added to the existing fuel locations until the flames from the piles reached the ceiling of the burn room again. Then, the shutters on the window were closed, and the temperature of the hot gas layer in the adjacent room was monitored in preparation for another gas cooling experiment.  The number of pallets added each time varied.

@@

The fuel load for test numbers 1 through 8 consisted of two sleeper sofas, two cushioned chairs, carpet, padding, and wood paneling along the southeast corner walls. An image of the fuel arrangement is shown in Figure~\ref{fig:Furniture_Fuel_Load}. The sofas and chairs were set up to mimic a typical seating area with one couch against the south wall and one couch against the east wall, each positioned 0.9~m (3.0~ft) away from the southeast corner. A chair was placed adjacent to each couch. The specific locations can be found in Figure~\ref{fig:Furniture_Fuel_Load_Dimensions}. In total, this fuel package contained approximately 336~kg (739~lb) of fuel comprised of wood, flexible polyurethane (PU), polyester (PE), and polypropylene (PP). Table~\ref{tab:Fire_Suppression_Fuel_Masses} provides a breakdown on the fuel composition for these experiments.

@@

The fuel load for test numbers 9 through 12 consisted of one bale of hay, 12 pallets, 11.5 sheets of plywood on the walls, and the wood ``floor assembly''. Two stacks of pallets with hay were the first items ignited in each experiment. Each stack was composed of six pallets with a half bale of hay layered between the six pallets. Figure~\ref{fig:Wood_Fuel_Load} shows the pallets prior to ignition. The pallets and hay were weighed before each experiment, and the average mass of the two stacks of pallets and the bale of hay over the four experiments was 232~kg (511~lb).

@@

The pallet stacks were located 0.3~m (1.0~ft) away from the east and south walls of the burn room (see Figure~\ref{fig:Wood_Fuel_Load_Dimensions}). The pallets were 1.2~m (3.9~ft) long, 1.0~m (3.3~ft) wide, and 123~mm (4.8~in) high. The stacks were spaced 150~mm (6.0~in) apart. One layer of 13~mm (0.5~in) thick gypsum board panels were laid on the concrete floor under the wood pallets to form a protective layer to minimize thermal damage to the concrete floor. The east and south walls of the burn room were covered with 15~mm (0.6~in) thick sheets of plywood. The ceiling of the burn room was composed of 12 wood TJIs and nine sheets of OSB that were part of the structure's ceiling support. The opening to the exterior on the south side of the burn room was covered with a 2.4~m (8.0~ft) tall by 1.2~m (4.0~ft) wide piece of medium density fiberboard paneling that was 5~mm (0.2~in) thick.

@@

In June of 2015, the larger compartment experiments, experiments 13 through 20, were conducted in the two-story structure.  Two different fuel packages located on the ground level were utilized throughout the experiments: experiments 13 and 14 used a fuel package composed primarily of furniture, and experiments 15 through 20 used the same fuel package with the addition of two stacks of wood pallets.

@@

The fuel load for experiments 13 and 14 consisted of three sofas with an average mass of 48.7~kg (107~lb); 51.4~m$^2$ (553~ft$^2$) of carpet and padding; and approximately 38 sheets of OSB paneling with dimensions 16~mm (0.6~in) x 1.2~m (4.0~ft) x 2.4~m (8.0~ft) along the east, west, and south walls and ceiling. The three couches were aligned along the west wall and spaced 1.2~m (4.0~ft) apart. The specific locations can be found in the dimensioned schematic Figure~\ref{fig:furniture_2story}. In total, this fuel package was composed of wood, flexible polyurethane (PU), polyester (PE), and polypropylene (PP).

@@

For experiments 15 through 20, two stacks of wood pallets were added to the fuel load consisting of three sofas; OSB lining the ceiling and walls; and the polypropylene carpeting over polyurethane foam padding. Each stack had 10 pallets and an approximate mass of 175~kg (386~lb). Hay from bales with an average a mass of 12~kg (26~lb) was added for ignition purposes. The specific locations of the fuel load's contents can be found in the dimensioned schematic Figure~\ref{fig:pallet_furniture_2story}.

@@

Spray density experiments were conducted by placing 0.23~m$^2$ (2.5~ft$^2$) interlocking water collection pans across the floor surface area of the burn rooms within each structure (see Figure~\ref{fig:water_buckets}). The depth of the pans was 0.3~m (1.0~ft).  Water or CAF was flowed for a specified duration or until it appeared that a pan was filled.  Upon completion of the test, the bins were removed from the structure. Then, each bin was weighed to quantify the impact of nozzle location and spray pattern. Thirty-three spray density tests were conducted: 21 in the burn building used for the gas cooling tests and 12 in the single-story structure used for the fire suppression experiments. The interior of the fire room in the single-story structure was sheathed in plastic for the spray density tests to protect the gypsum board from being soaked with water prior to the fire suppression experiments. The flow rate for each test was 120~gpm for water and 120~gpm plus 60~cfm air for CAF. Tables~\ref{tab:spray_density_tests} and \ref{tab:spray_density_tests2} provide an overview of each experiment. In the table, ``GC'' stands for the gas cooling experiments, and ``SE'' stands for the suppression experiments. For all experiments, 30.0~m (100~ft) of 45~mm (1.75~in) diameter hoseline connected to a ``blitz-fire'' monitor with a Metro 1 nozzle was used.

@@

For the spray density tests listed in Table~\ref{tab:spray_density_tests}, the ``mid'' position had the centerline of the stream aimed at 2.4~m (8.0~ft) south of the north wall and 2.1~m (7.0~ft) west of the east wall in the concrete burn building. The ``back'' position had the centerline of the stream aimed at 3.4~m (11.0~ft) south of the north wall and 2.1~m (7.0~ft) west of the east wall in the concrete burn building. Figure~\ref{fig:stream_position} shows the streams from the monitor nozzle for the middle and back positions. From the images, the qualitative difference of the nozzle positions can be observed.  When the stream is in the middle position, after the stream hits the ceiling, the impact region is relatively narrow with the majority of the water hitting the walls of the corner opposite the nozzle position.  When the nozzle is in the back position, the stream hits the ceiling at a much steeper angle and the droplets from the broken stream have a broader distribution throughout the room.

@@

\begin{table}[!ht]
\footnotesize
\centering
\captionof{table}{Spray Density Test Matrix for Gas Cooling Experiments}\label{tab:spray_density_tests}
\begin{tabular}{lllll}
\toprule[1.5pt]
Test \#    &  Duration (s)  & Agent  &  Pattern            	& Nozzle Position  \\
\midrule
GC1        &  15            & Water  &  Straight Stream    	& Back             \\
GC2        &  23            & Water  &  Straight Stream    	& Back             \\
GC3        &  15            & Water  &  Straight Stream    	& Mid              \\
GC4        &  15            & Water  &  Straight Stream    	& Mid              \\
GC5        &  20            & Water  &  Straight Stream 	& Mid              \\
GC6        &  21            & Water  &  Narrow Fog   		& Mid              \\
GC7        &  20            & Water  &  Narrow Fog   		& Mid              \\
GC8        &  22            & Water  &  Narrow Fog   		& Back             \\
GC9        &  24            & Water  &  Narrow Fog   		& Back             \\
GC10       &  20            & Water  &  Solid Stream       	& Back             \\
GC11       &  22            & Water  &  Solid Stream       	& Back             \\
GC12       &  15            & CAF    &  Solid Stream       	& Back             \\
GC13       &  10            & CAF    &  Solid Stream       	& Mid              \\
GC14       &  15            & CAF    &  Solid Stream       	& Back             \\
GC15       &  20            & CAF    &  Solid Stream       	& Mid              \\
GC16       &  15            & CAF    &  Narrow Fog   		& Mid              \\
GC17       &  20            & CAF    &  Straight Stream    	& Mid             \\
GC18       &  10            & CAF    &  Straight Stream    	& Mid             \\
GC19       &  15            & CAF    &  Narrow Fog   		& Back             \\
GC20       &  15            & CAF    &  Narrow Fog   		& Back             \\
GC21       &  15            & CAF    &  Straight Stream    	& Back             \\
\bottomrule[1.25pt]
\end{tabular}\par
\end{table}
For the fire suppression spray density experiments listed in Table~\ref{tab:spray_density_tests2}, the ``hallway'' position had the center of the stream aimed at the ceiling of the burn room, 2.1~m (7.0~ft) north of the south wall and 1.5~m (5.0~ft) east of the west wall. The ``fire room'' position had the center of the stream aimed at the ceiling of the burn room, 2.1~m (7.0~ft) north of the south wall and 3.1~m (10.0~ft) west of the east wall.

@@

To further analyze the data, the temperature differences can be separated by each of the three thermocouple arrays in the gas cooling room. Figure~\ref{fig:Gas_Cooling_Instrumentation_Dimensions} shows the placement of the three thermocouple arrays (TC A1, TC A2, and TC A3). Ideally, a thermocouple provides an estimate of the temperature of the gas surrounding it. However, a thermocouple actually measures its own temperature. So, if a droplet of water or some CAF clings to a thermocouple, the temperature provided by that thermocouple will be different than the surrounding gas temperature.
